News Across Languages - Cross-Lingual Document Similarity and Event Tracking

### 2015-09-04 rupnik2015.tex
related work

\begin{figure}
\centering
\includegraphics[width=1\textwidth]{events2.png}
\caption{\label{fig:event2} Events are represented by collections of articles about an event, in this case the Malaysian airliner which was shot down over Ukraine. The results shown in the figure can be obtained using the query \url{http://eventregistry.org/event/997350\#?lang=eng\&tab=articles}.}
\end{figure}

@@

Our goal is to build a system that monitors global media and analyzes how events are being reported on. Our approach consists of two steps: tracking events separately in each language (based on language detection and an online clustering approach) and then connecting them. The pipeline must be able to process millions of articles per day and perform billions of similarity computations each day. Both step rely heavily on simliarity computation, which implies that this must be highly scalable.
%
%while both steps must be scalable, the underlying similarity computation is used
%- implying that similarity computation should be highly scalable.  %The system should support as many languages as possible.
Therefore, we focus on implementations that run on a single shared memory machine, as opposed to clusters of machines. This  simplifies implementation and system maintenance.
To summarize, the following properties are desirable:
\begin{itemize}
\item \textbf{Training} - The training (building cross-lingual models) should scale to many languages and should be robust to the quality of training resources. The system should be able to take advantage of comparable corpora (as opposed to parallel translation-based corpora), with missing data. Supporting a new language should be straightforward.
\item \textbf{Operation efficiency} - The similarity computation should be fast - the system must be able to handle billions of similarity computations per day.
\item \textbf{Operation cost} - The system should run on a strong shared machine server and not rely on paid services.
\item \textbf{Implementation} - The system is straightforward to implement, with few parameters to tune.
\end{itemize}
We believe that a cross-lingual similarity component that meets such requirements is very desirable in a commercial setting, where several different costs have to be taken into consideration.

@@

\textbf{Translation}. The most obvious way to compare documents written in different languages is to use machine translation and perform monolingual similarity. One can use free tools such as Moses \shortcite{moses} or translation services, such as Google Translate\footnote{https://translate.google.com/}. There are two issues with such approaches: they solve a harder problem than needs to be solved and they are less robust to training resource quality - large sets of translated sentences are typically needed. Training Moses for languages with scarce linguistic resources is thus problematic. The issue with using online services such as Google Translate is that the APIs are limited and not free. The operation efficiency and cost requirements make translation-based approaches less suited for our system.

@@

\textbf{Matrix factorization}. Non-negative matrix factorization based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CL-LSI\shortcite{cl_lsi}\cite{multilingualBook}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices of dimensions that is the product of training set size and vocabulary size). Our proposed approach combines CCA and CL-LSI. Another closely related method is Explicit Semantic Analysis (ESA) \cite{ESA}, which uses Wikipedia (as do we in the current work) to compare documents. The authors of ESA compare it to CL-LSI and find that CL-LSI can outperform ESA in an information retrieval, but is costlier to optimize over a large corpus (ESA requires no training). We find that the scalability argument does not apply in our case: based on advances in numerical linear algebra we can solve large CL-LSI problems (millions of documents as opposed to 10,000 document limit reported in \cite{ESA}).

>>>>

\begin{figure}
\centering
\includegraphics[width=1\textwidth]{events2.png}
\caption{\label{fig:event2} Events are represented by collections of articles about an event, in this case the Malaysian airliner which was shot down over Ukraine. The results shown in the figure can be obtained using the query \url{http://eventregistry.org/event/997350\#?lang=eng\&tab=articles}. The content presented is part of the Event Registry system, developed by the authors.}
\end{figure}

@@

Our goal is to build a system that monitors global media and analyzes how events are being reported on. Our approach consists of two steps: tracking events separately in each language (based on language detection and an online clustering approach) and then connecting them. The pipeline must be able to process millions of articles per day and perform billions of similarity computations each day. Both step rely heavily on simliarity computation, which implies that this must be highly scalable.
%
%while both steps must be scalable, the underlying similarity computation is used
%- implying that similarity computation should be highly scalable.  %The system should support as many languages as possible.
Therefore, we focus on implementations that run on a single shared memory machine, as opposed to clusters of machines. This  simplifies implementation and system maintenance.
To summarize, the following properties are desirable:
\begin{itemize}
\item \textbf{Training} - The training (building cross-lingual models) should scale to many languages and should be robust to the quality of training resources. The system should be able to take advantage of comparable corpora (as opposed to parallel translation-based corpora), with missing data. Supporting a new language should be straightforward.
\item \textbf{Operation efficiency} - The similarity computation should be fast - the system must be able to handle billions of similarity computations per day. Computing the similarity between a new document and a set of known documents should be efficient (the main application is linking documents between two monolingual streams).
\item \textbf{Operation cost} - The system should run on a strong shared machine server and not rely on paid services.
\item \textbf{Implementation} - The system is straightforward to implement, with few parameters to tune.
\end{itemize}
We believe that a cross-lingual similarity component that meets such requirements is very desirable in a commercial setting, where several different costs have to be taken into consideration.

@@

\textbf{Translation and dictionary based}. The most obvious way to compare documents written in different languages is to use machine translation and perform monolingual similarity, see  \cite{multilingualBook}\cite{plagiarism} for several variations of translation based approaches. One can use free tools such as Moses \shortcite{moses} or translation services, such as Google Translate\footnote{https://translate.google.com/}. There are two issues with such approaches: they solve a harder problem than needs to be solved and they are less robust to training resource quality - large sets of translated sentences are typically needed. Training Moses for languages with scarce linguistic resources is thus problematic. The issue with using online services such as Google Translate is that the APIs are limited and not free. The operation efficiency and cost requirements make translation-based approaches less suited for our system. CL-VSM \cite{plagiarism} and the approach presented in \cite{pouliquen2008story} compare documents by using dictionaries, such as EuroVoc \cite{eurovoc}. The generality of such approaches is limited by the quality of available linguistic resources, which may be scarce or non-existent for certain language pairs.

@@

\textbf{Matrix factorization}. Non-negative matrix factorization based \cite{nonnegfactor_lsi}, Cross-Lingual Latent Semantic Indexing CL-LSI\shortcite{cl_lsi}\cite{multilingualBook}, Canonical Correlation Analysis (CCA) \cite{Hotelling}, Oriented Principal Component Analysis \cite{platt2010translingual}. The quadratic time and space dependency of the OPCA method makes it impractical for large scale purposes. In addition, OPCA forces the vocabulary sizes for all languages to be the same, which is less intuitive. For our setting, the method in \cite{nonnegfactor_lsi} has a prohibitively high computational cost when building models (it uses dense matrices of dimensions that is the product of training set size and vocabulary size). Our proposed approach combines CCA and CL-LSI. Another closely related method is Explicit Semantic Analysis (ESA) \cite{ESA}, which uses Wikipedia (as do we in the current work) to compare documents. It can be interpreted as using the sample covariance matrix between features of two languages to define the dot product that is used to compute similarities. 
The authors of ESA compare it to CL-LSI and find that CL-LSI can outperform ESA in an information retrieval, but is costlier to optimize over a large corpus (ESA requires no training). We find that the scalability argument does not apply in our case: based on advances in numerical linear algebra we can solve large CL-LSI problems (millions of documents as opposed to 10,000 document limit reported in \cite{ESA}). In addition, ESA is less suited for computing similarities between two large monolingual streams. For example, each day we have to compute similarities between 500,000 English and 500,000 German news articles. Comparing each German news article with 500,000 English news articles is either prohibitively slow (involves projecting all English articles on Wikipedia) or consumes too much memory (involves storing the projected English articles, which for a Wikipedia of size 1,000,000 is a 500,000 by 1000,0000 non-sparse matrix).

\textbf{Monolingual}. Finally, related work includes monolingual approaches that treat document written in different languages in a monolingual fashion. The intuition is that named entities (for example, ``Obama'') and cognate words (for example, ``tsunami'') are written in the same or similar fashion in many languages. For example, the Cross-Language Character n-Gram Model (CL-CNG) \cite{plagiarism} represents documents as bags of character n-grams. Another approach is to use language dependent keyword lists based on cognate words \cite{pouliquen2008story}. These approaches may be suitable for comparing documents written in languages that share a writing system, which does not apply to the case of global news tracking.
